Water desalination using polyelectrolyte hydrogel: Gibbs ensemble modeling

### 2022-09-15 gibbs.tex
Update gibbs.tex

Recently, Li \etal \cite{Li2011} took advantage of the smart hydrogel properties for desalination purposes. In this work, the authors used hydrogels as draw agents for FO. 
They have demonstrated that the hydrogels are able to absorb water across the FO membrane due to their swelling and osmotic pressure and
allow dewatering under the influence of stimuli (thermal and/or mechanical), due to their insoluble cross-linked polymer network.
They have proposed the use of hydrogel based on thermoresponsive polyelectrolyte~--- copolymer of poly-N-isopropyl acrylamide (p-NIPAAm) and polyacrylic acid (p-AA).
This gel network is either hydrophilic or hydrophobic, depending on the temperature.
Such gel accumulates water inside the network in its hydrophilic state, but it releases the water out in its hydrophobic state.

@@

The lower salinity solution can be extracted from the gel by means of compression and/or other stimulus providing that the charge of the gel remains constant.
In the case of gel made of weak (pH-sensitive) polyelectrolyte, the compression discharges the gel, therefore the neutralizing counterions leave out the gel diminishing the desalination effect \cite{Rud2018}. 
%Nevertheless, in\cite{Rud2018} we discussed how the release of counterions may be employed for desalination as well.

One can argue that the sole Donnan effect is insufficient to achieve a high salt rejection \cite{Cai2016} and the salinity of the water squeezed from hydrogel under a very high hydraulic pressure (up to 100 bar \cite{Fengler_2020}) turns out to be not much different from the initial.
Indeed, the use of high hydraulic pressure diminishes all the advantages of this method over RO, and the reversibility of hydrogel swelling after strong compression remains questionable. 
Nevertheless, in the current study, we model the compression of the gel limiting ourselves to low compression rates, less than 5 bar, and studied how the compression of the gel affects the surrounding salinity.
We modelled also the desalination process, as a cascade of following one by each other gel swellings and compressions driving the salinity of supernate down to drinkable water. 
%In the current study we would like to concentrate on the use of gel made of strong polyelectrolyte for water desalination. 
%Namely, to investigate how the compression of the gel affects the salinity of the surrounding the gel solution and to propose the way of water desalination employing the hydrogel swelling/deswelling set of processes.

%The target of our study is to model the thermodynamic equilibrium between the two boxes and to show how the compression of the gel affects the ion density in the outside solution.

TODO Then, highlight the water desalination technologies and Gibbs ensemble (https://doi.org/10.3390/nano10020293 \cite{erdos2020}).

>>>>

Li \etal \cite{Li2011} took advantage of the smart hydrogel properties for desalination purposes. In this work, the authors used hydrogels as draw agents for FO. 
They have demonstrated that the hydrogels are able to absorb water across the FO membrane due to their swelling and osmotic pressure and
allow dewatering under the influence of stimuli (thermal and/or mechanical), due to their insoluble cross-linked polymer network.
They have proposed the use of hydrogel based on thermoresponsive polyelectrolyte~--- copolymer of poly-N-isopropyl acrylamide (p-NIPAAm) and polyacrylic acid (p-AA).
This gel network is either hydrophilic or hydrophobic, depending on the temperature.
Such gel accumulates water inside the network in its hydrophilic state, but it releases the water out in its hydrophobic state.

@@

As shown in \cite{Li_2011,Ali_2015,Arens_2017,Fengler_2020}, the solution of lower salinity can be extracted from the gel by means of compression and/or other stimulus, provided that the charge in the gel remains constant.
In the case of weak polyelectrolyte (pH-sensitive) gel, the compression discharges the gel and therefore the neutralizing counterions leave out the gel, diminishing the desalination effect \cite{Rud2018}.

One can argue that the sole Donnan effect is insufficient to achieve a high salt rejection \cite{Cai2016} and the salinity of the water squeezed from hydrogel under a very high hydraulic pressure (up to 100 bar \cite{Fengler_2020}) turns out to be not much different from the initial.
Indeed, the use of high hydraulic pressure diminishes all the advantages of this method over RO, and the reversibility of hydrogel swelling after strong compression remains questionable. 
Nevertheless, in this study, we model the compression of the gel limiting ourselves to low compression rates, less than 5 bar, and studied how the compression of the gel affects the surrounding salinity.
We modelled the desalination process, as a cascade of step-by-step gel swellings and compressions, driving the salinity of the supernate down to potable water.

% water desalination technologies and Gibbs ensemble (https://doi.org/10.3390/nano10020293 \cite{erdos2020}).
